An algorithm for finding the global extremum\\ of a partially defined function\thanks{This work was supported by the Russian Science Foundation, project No. 21-11-00204.}

### 2023-10-04 2023/MMST/samplepaper.tex
last fix after first review

Third, the objective function satisfies the Lipschitz condition
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
where $L$ is the Lipschitz constant. The dimensionality reduction scheme using Peano curves is known to juxtapose the multidimensional problem with a Lipschitz objective function (\ref{eq1}) to problem (\ref{eq2}) with the one-dimensional objective function satisfying H{\"o}lder condition
\begin{equation}\label{eq4} 
| \phi (x')-\phi (x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
\end{equation}
where $\rho(x',x'') =  |x' - x''|^{1/N}$ and $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.

@@

\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 0 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
\end{equation}
and juxtapose them with the values $z_i=\phi (x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
    1  & {\quad \text{if } x_i \text{ is an internal point.}}
  \end{cases}
\end{equation}
The points $x_0=0$ and $x_{k}=1$ are introduced additionally (the values $z_0$ and $z_{k}$ are undefined) for convenience of the following notations.

\paragraph{Rule 2.} Compute the estimate
\begin{equation}\label{eq7} 
\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= (x_i-x_{i-1})^{1/N},
\end{equation}
of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $\phi(x)$, where $r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
\paragraph{Rule 3.} Determine the best current value of the objective function
\begin{equation}\label{eq8} 
z^*=\min \left\{ \phi (x_i): 1\leq i \leq k \right\}.
\end{equation}

@@

\subsubsection{Description of the modified algorithm.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
  \item The value $z_v^*$ depends on the index value
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ \phi (x_i): v(x_i)=1, 1 \leq i \leq k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r > 0$ is a parameter of the method.

>>>>

Third, the objective function satisfies the Lipschitz condition
\begin{equation}\label{eq3} 
| \phi (y')-\phi (y'') | \leq L \| y'-y'' \|, \; y',y'' \in D,
\end{equation}
where $L$ is the Lipschitz constant. The dimensionality reduction scheme using Peano curves is known to juxtapose the multidimensional problem with a Lipschitz objective function (\ref{eq1}) to problem (\ref{eq2}) with the one-dimensional objective function satisfying H{\"o}lder condition
\begin{equation}\label{eq4} 
| f(x')-f(x'') | \leq K \rho(x',x''), \; x',x'' \in [0,1],
\end{equation}
where $\rho(x',x'') =  |x' - x''|^{1/N}$ and $N$ is the dimensionality of the initial multidimensional problem and coefficient $K$ is related to the Lipschitz constant $L$ by the relation $K \leq 2L\sqrt {N+3}$ \cite{Strongin2000}.

@@

\paragraph{Rule 1.} Renumber (by the lower indices) the points $x^i, 0 \leq i \leq k$  of the previous trials in the ascending order by their coordinates, i.e.
\begin{equation}\label{eq5} 
0=x_0 < x_1 < ... < x_i < ... < x_{k}=1
\end{equation}
and juxtapose them with the values $z_i=f(x_i), 0 < i < k$, computed at these points and to the indices $v_i=v(x_i)$ defined by the rule
\begin{equation}\label{eq6} 
v_i=v(x_i)=
  \begin{cases}
    -1 & {\quad \text{if } x_i \text{ is a boundary point,}}\\
    1  & {\quad \text{if } x_i \text{ is an internal point.}}
  \end{cases}
\end{equation}
The points $x_0=0$ and $x_{k}=1$ are introduced additionally (the values $z_0$ and $z_{k}$ are undefined) for convenience of the following notations.

\paragraph{Rule 2.} Compute the estimate
\begin{equation}\label{eq7} 
\mu = \max\left\{ \frac{|z_i-z_{i-1}|}{\Delta _i}, 2 \leq i \leq k \right\} , \Delta _i= (x_i-x_{i-1})^{1/N},
\end{equation}
of the unknown H{\"o}lder constant $K$ from (\ref{eq4}) for the reduced function $f(x)$, where $r>1$ is a parameter of the algorithm; if $\mu=0$ accept $\mu=1$.
\paragraph{Rule 3.} Determine the best current value of the objective function
\begin{equation}\label{eq8} 
z^*=\min \left\{ f(x_i): 1\leq i \leq k \right\}.
\end{equation}

@@

\subsubsection{Description of the modified algorithm.} The algorithm remains the same with an introduction of the following minor additions in the computing formulae in the case of falling into the non-computable search regions.
\begin{enumerate} 
  \item The index $v_i=v(x_i)$ is determined by the formula (\ref{eq12}).
  \item When computing current lower estimate of the H{\"o}lder constant $K$ in formula (\ref{eq7}), use only the intervals with two computable boundary points.
  \item The value $z_v^*$ depends on the index value
\begin{equation}\label{eq20} 
z_v^*=
  \begin{cases}
    -\varepsilon _r, & {\quad v=0 ,}\\
    \min \left\{ f(x_i): v(x_i)=1, 0 < i < k \right\}, & {\quad v=1 ,}
  \end{cases}
\end{equation}
where $\varepsilon _r > 0$ is a parameter of the method.
